### 2022-11-03 3rd_try.qmd
Create 3rd_try.qmd

---
title: "quato_file_2"
format: html
---

## revised Quarto file

This is a new try of generating a Quarto file in RStudio. The engine is revised as Knitr. Below is a try-out description of the corpus.

BROWN CORPUS

A Standard Corpus of Present-Day Edited American

English, for use with Digital Computers.

by W. N. Francis and H. Kucera (1964)

Department of Linguistics, Brown University

Providence, Rhode Island, USA

Revised 1971, Revised and Amplified 1979

http://www.hit.uib.no/icame/brown/bcm.html

Distributed with the permission of the copyright holder,

redistribution permitted.