### 2023-07-18 practice with ggplot.qmd
updated paper version

---
title: "ggplot_tryout"
format: html
code-fold: true
---

```{r}
library(tidyverse)
set.seed(2022)
rows <- 20
cake_eaten <- tibble(
  hunger_level = sample(1:100, rows, replace = TRUE),
  cake_eaten = sample(1:100, rows, replace = TRUE)/100, 
  cake_flavor = sample(rep_len(c("vanilla", "chocolate"), rows))
)
```

##Step1: setting up a plot

```{r}
#| label: fig-ggplot
#| fig-cap: "Output of the basic `ggplot()` call."
g <- ggplot(cake_eaten, aes(x = cake_eaten, y = hunger_level))
g
```

##Step2: adding fig points

```{r}
#| label: fig-point
#| fig-cap: Output after adding the points geometry.
#| code-fold: true
g <- g + geom_point(aes(color = cake_flavor), size = 3)
g
```

##Step 3: adding fig-theme-light

```{r}
#| label: fig-theme-light
#| fig-cap: Output after adding a "light" theme.
g <- g + theme_light(base_size = 20)
g
```